Splitting It Up: The \pkg{spduration} Split-Population Duration Regression Package for Time-varying Covariates

### 2017-09-26 R Journal/beger-hill-metternich-minhas-ward.tex
expand Table 1, fixes #3

\begin{table}
\begin{center}
\begin{tabular}{lrrrrr} \toprule
Country & Year & Coup & Spell ID & Duration \\ \midrule
Portugal & 1960 & 0 & 1 & 1  \\ 
\ldots & \ldots & 0 & 1 & \ldots \\
\ldots & 1975 & 1 & 1 & 16 \\ \midrule
Portugal & 1976 & 0 & 2 & 1 \\
\ldots & \ldots & 0 & 2 & \ldots \\
\ldots & 2000 & 0 & 2 & 20  \\ \midrule
Canada & 1960 & 0 & 3 & 1 \\
\ldots & \ldots & 0 & 3 & \ldots \\ 
\ldots & 2000 & 0 & 3 & 41 \\ \bottomrule
\end{tabular}
\end{center}
\caption{Example data frame with select duration variables.} \label{tab-ex}
\end{table}

The \code{add\_duration} functions identifies spells in the data and
returns the data frame with several additional variables needed for
estimation. The most important among these are a counter of elapsed time
in a spell, \code{duration}. It also codes that a spell was at risk
(\code{atrisk==1}) if it ended in a failure event, and not at risk
otherwise. These are then used as response variables for the formula in
the \code{spdur} function.

>>>>

\begin{table}
\begin{center}
\begin{tabular}{lrrrrrrr} \toprule
Country & Year & Coup & Spell ID & Failure & Censor & At risk & Duration \\ \midrule
Portugal & 1960   & 0 & 1 & 0 & 0 & 1 & 1  \\ 
\ldots   & \ldots & 0 & 1 & 0 & 0 & 1 & \ldots \\
\ldots   & 1975   & 1 & 1 & 1 & 0 & 1 & 16 \\ \midrule
Portugal & 1976   & 0 & 2 & 0 & 0 & 0 & 1 \\
\ldots   & \ldots & 0 & 2 & 0 & 0 & 0 & \ldots \\
\ldots   & 2000   & 0 & 2 & 0 & 1 & 0 & 20  \\ \midrule
Canada   & 1960   & 0 & 3 & 0 & 0 & 0 & 1 \\
\ldots   & \ldots & 0 & 3 & 0 & 0 & 0 & \ldots \\ 
\ldots   & 2000   & 0 & 3 & 0 & 1 & 0 & 41 \\ \bottomrule
\end{tabular}
\end{center}
\caption{Example data frame with select duration variables.} \label{tab-ex}
\end{table}

The \code{add\_duration} functions identifies spells in the data and
returns the data frame with several additional variables needed for
estimation. The most important of these are shown in Table \ref{tab-ex}. Using coups as the outcome, failure is coded when a coup occured. This also triggers a new spell if there are more observations for a country. If a spell ends without a failure event it is coded as right censored (\code{censor}). This can occur either because we have reached the end of observed data--the year 2000, or because a country dropped out of the data, e.g. if it was subsumed by another country. Duration is then coded each spell by counting the number of time periods until a spell ends. The at risk coding depends on whether a spell ended in failure--spells that end in failure are coded as being at risk for their entire duration, otherwise they are coded as not being at risk.

The \code{add\_duration} function also adds several other variables that should be of less concern in typical use. If the data contain outcomes that can occur over multiple time periods, like civil wars, \code{add\_duration} should be used with the \code{ongoing = TRUE} flag so that only the first time period is treated as a failure, rather than as repeated failures. In that case \code{ongoing} is a marker for ongoing events beyond the initial onset. The variable \code{cured} is the inverse of \code{atrisk}, i.e. $1 -$~\code{atrisk}. Lastly, \code{t.0} marks the starting time for the row. It is by default \code{duration}~$ - 1$, but is retained to allow extension, if needed in the future, for observations that do not span a fixed time period, e.g. if observations are taken at irregular time intervals.
